---
title: "NMR Visits"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "I:/Documents/PhD files/Kim/UPBEAT/Data Files")
```

As the NMR data comes in .dta format, we need to read in the library 'haven' to be able to read the data format in R.

```{r}
library(haven)
setwd("I:\\Science\\MS\\users\\students\\Dygas_Natalia\\Documents\\PhD files\\Kim\\UPBEAT\\Data Files")
meta <- read_dta("upbeat_metabolite_biomarkers_all.dta") #full metabolome data

```


Now, we need to isolate the participant_id values which have metabolite data for visit 1, 3 and 5.

This creates a data set of participant_id values which have metabolite data for visit 1, 3 and 5, but also keeps all other visits. Need to filter out the unnecessary data of followup values.

```{r}
meta %>% group_by(participant_id) %>% 
  filter("Visit 1" %in% visno & "Visit 3" %in% visno & "Visit 5" %in% visno, visno %in% c("Visit 1","Visit 3","Visit 5")) %>% 
  ungroup() -> meta_visits

summary(as.factor(meta_visits$visit))
```
From this we can see that 712 participants have metabolomic data for all 3 visits. Now, we need to link this to the obesity data, and see how many of these 712 praticipants have a binary outcome for obesity/overweight(+).

```{r}
setwd("I:\\Science\\MS\\users\\students\\Dygas_Natalia\\Documents\\PhD files\\Kim\\UPBEAT\\Data Files")
obesity <- read.csv("obesity_child_cat.csv")
```


```{r}
bind_meta <- merge(meta_visits, obesity, by="participant_id")
length(unique(bind_meta$participant_id))
```
```{r}
setwd("I:\\Science\\MS\\users\\students\\Dygas_Natalia\\Documents\\PhD files\\Kim\\UPBEAT\\Data Files")
write.csv(bind_meta, "metabol_raw_data.csv", row.names= F)
```



In this data, we have 266 unique participants, i.e., there are 266 participants with metabolomic data for all 3 visits which have a binary outcome for obesity/overweight(+).

Need to create a dataset which includes the participant_id, f72_neo_sex, category, child_adip_over, and all metabolites from the NMR metabolite data set, but looking at the rate of change per metabolite between visits i.e., visit 1 to visit 3, visit 3 to visit 5 and visit 1 to visit 5. 

```{r}
bind_meta[,c(1,4,6:233, 260, 261, 264:266)] %>%
  arrange(participant_id,visit) %>%
  mutate(change=case_when(visit==1~"1-3",visit==3~"3-5",visit==5~"1-5")) %>%
  group_by(participant_id) %>%
  mutate_at(vars(-c(participant_id, visit, f72_neo_sex.y, mbodpodbmi, Category, child_adip_over, child_adip_ob, change)),
            ~case_when(visit==1~(nth(.,2)-first(.))/first(.)*100,
                       visit==3~(nth(.,3)-nth(.,2))/nth(.,2)*100,
                       visit==5~(nth(.,3)-first(.))/first(.)*100)) %>%
  relocate(change,.before="visit") %>%
  select(-visit) -> bind_meta_change

```

```{r}
bind_meta_change[bind_meta_change=="Inf"] <- NA
bind_meta_change$change <- as.factor(bind_meta_change$change)
```

The new dataset bind_meta_change shows the rates of change in % for each metabolite. The visits from which the numbers are taken from are mentioned in the 'change' variable. A negative value shows a % decrease, and a positive value shows a % increase i.e. -50.945 is a 50.945% decrease, and 162.85 is a 62.85% increase in metabolite volume between visits. The NAs came about as some of the metabolites had values of 0 before calculating rates of change, which generated Inf values.  

We will now save the new data set with rates of change as a new data set.

```{r}
setwd("I:\\Science\\MS\\users\\students\\Dygas_Natalia\\Documents\\PhD files\\Kim\\UPBEAT\\Data Files")
write.csv(bind_meta_change, "metabol_rates_of_change.csv", row.names= F)
```




